$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-05-17 paper/e6ssm-tuning-paper.tex
simplified some of the statements I added.

$U(1)$ extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{Fayet:1977yc, Kim:1983dt, Suematsu:1994qm,
  Cvetic:1995rj, Cvetic:1996mf, Jain:1995cb, Nir:1995bu,
  Cvetic:1997ky} and also raise the Higgs mass with new $F$- and
$D$-terms.  Nonetheless, as was recently demonstrated in the context
of the exceptional supersymmetric standard model (E$_6$SSM)
\cite{King:2005jy, King:2005my, Athron:2010zz}, such models can still
suffer from naturalness problems with the mass of the new $Z^\prime$
associated with the break down of the new $U(1)$ appearing in the
electroweak symmetry breaking conditions at tree level
\cite{Athron:2013ipa}.  Despite this the constrained version of the
E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue, Athron:2009bs} was still
found to be significantly less tuned than the cMSSM.  It has recently
been pointed out to us that this tree level fine tuning effect from
the $Z^\prime$ mass was also considered previously \cite{Drees:1985js}.

@@

The extra $U(1)^\prime$ gauge symmetry may mix with the $U(1)_Y$ gauge
symmetry associated with hypercharge through gauge kinetic mixing,
\be
\mathcal{L}_{mix}^{kin}=-\frac{\sin\chi}{2}F^{Y}_{\mu\nu}F^{N}_{\mu\nu}\,.
\label{Eq:GKM}
\ee
where $F_{\mu\nu}^Y$ and $F_{\mu\nu}^{N}$ are field strengths associated
with the $U(1)_Y$ and $U(1)_{N}$ respectively.  The gauge kinetic mixing
can have a significant impact on the phenomenology
\cite{Rizzo:1998ut,Salvioni:2009mt,Krauss:2012ku}, which may in particular
reduce the $Z^\prime$ limit mass by up to $200$ GeV.

@@

Using the approach outlined above, we are able to scan the low energy
parameter space of the MSSM and E$_6$SSM and calculate the fine tuning
in each.  To do so, we implemented the above expressions for computing
the fine tuning in a modified version of the E$_6$SSM spectrum
generator that was used in \cite{Athron:2013ipa}.  This code
implemented two-loop RGEs for all parameters except the soft scalar
masses.  In order to properly include the fine tuning impact of the
$SU(3)$ gaugino soft mass $M_3$, we have extended the original code to
make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,
  Staub:2010jh, Staub:2012pb, Staub:2013tta} and FlexibleSUSY
\cite{Athron:2014yba}, which also makes use of SOFTSUSY
\cite{Allanach:2001kg, Allanach:2013kza}.  The CP-even Higgs masses
are calculated including the leading one-loop effective potential
contributions given in \cite{Athron:2009bs} and for the light Higgs we
use the leading two-loop\footnote{Two-loop corrections calculated for
a non-minimal SUSY model may now also be obtained from SARAH
\cite{Goodsell:2014pla,Goodsell:2015ira}.  However this was not part
of SARAH when we were doing the numerical work in this project and
not currently used in FlexibleSUSY.  Such corrections go beyond the
required precision for studying fine tuning here.} contributions from
\cite{King:2005jy} which are a generalization of the corrections in the
MSSM and NMSSM calculated using effective field theory techniques
\cite{Carena:1995wu, Ellwanger:1999ji}.  To scan over the MSSM
parameter space, the equivalent MSSM fine tuning expressions were
implemented into a modified version of SOFTSUSY 3.3.10
\cite{Allanach:2001kg}.  For consistency with the results produced in
the $E_6$ models, and for computational speed, for our main scans only
the dominant one- and two-loop corrections to the CP-even Higgs masses
were included.  Finally, in all of the results below the fine tuning
was evaluated at the scale $Q = M_{\textrm{SUSY}} =
\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$
are the running $\overline{\textrm{DR}}$ stop masses evaluated at $Q =
M_{\textrm{SUSY}}$.

>>>>

$U(1)$ extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{Fayet:1977yc, Kim:1983dt, Suematsu:1994qm,
  Cvetic:1995rj, Cvetic:1996mf, Jain:1995cb, Nir:1995bu,
  Cvetic:1997ky} and also raise the Higgs mass with new $F$- and
$D$-terms.  Nonetheless, as was recently demonstrated in the context
of the exceptional supersymmetric standard model (E$_6$SSM)
\cite{King:2005jy, King:2005my, Athron:2010zz}, such models can still
suffer from naturalness problems with the mass of the new $Z^\prime$
associated with the break down of the new $U(1)$ appearing in the
electroweak symmetry breaking conditions at tree level
\cite{Athron:2013ipa}.  Despite this the constrained version of the
E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue, Athron:2009bs} was still
found to be significantly less tuned than the cMSSM.  Tree level
fine tuning from the $Z^\prime$ mass was also considered
previously \cite{Drees:1985js}.

@@

The extra $U(1)^\prime$ gauge symmetry may mix with the $U(1)_Y$ gauge
symmetry associated with hypercharge through gauge kinetic mixing,
\be
\mathcal{L}_{mix}^{kin}=-\frac{\sin\chi}{2}F^{Y}_{\mu\nu}F^{N}_{\mu\nu}\,.
\label{Eq:GKM}
\ee where $F_{\mu\nu}^Y$ and $F_{\mu\nu}^{N}$ are field strengths
associated with the $U(1)_Y$ and $U(1)_{N}$ respectively.  The gauge
kinetic mixing can have a significant impact on the phenomenology
\cite{Rizzo:1998ut,Salvioni:2009mt,Krauss:2012ku}, and may reduce the
$Z^\prime$ mass limit.

@@

Using the approach outlined above, we are able to scan the low energy
parameter space of the MSSM and E$_6$SSM and calculate the fine tuning
in each.  To do so, we implemented the above expressions for computing
the fine tuning in a modified version of the E$_6$SSM spectrum
generator that was used in \cite{Athron:2013ipa}.  This code
implemented two-loop RGEs for all parameters except the soft scalar
masses.  In order to properly include the fine tuning impact of the
$SU(3)$ gaugino soft mass $M_3$, we have extended the original code to
make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,
  Staub:2010jh, Staub:2012pb, Staub:2013tta} and FlexibleSUSY
\cite{Athron:2014yba}, which also makes use of SOFTSUSY
\cite{Allanach:2001kg, Allanach:2013kza}.  The CP-even Higgs masses
are calculated including the leading one-loop effective potential
contributions given in \cite{Athron:2009bs} and for the light Higgs we
use the leading two-loop\footnote{Two-loop corrections calculated for
  a non-minimal SUSY model may now also be obtained from SARAH
  \cite{Goodsell:2014pla,Goodsell:2015ira}.  However this was not
  available when the numerical work for this paper was carried out and
  such corrections go beyond the required precision for studying fine
  tuning here.} contributions from \cite{King:2005jy} which are a
generalization of the corrections in the MSSM and NMSSM calculated
using effective field theory techniques \cite{Carena:1995wu,
  Ellwanger:1999ji}.  To scan over the MSSM parameter space, the
equivalent MSSM fine tuning expressions were implemented into a
modified version of SOFTSUSY 3.3.10 \cite{Allanach:2001kg}.  For
consistency with the results produced in the $E_6$ models, and for
computational speed, for our main scans only the dominant one- and
two-loop corrections to the CP-even Higgs masses were included.
Finally, in all of the results below the fine tuning was evaluated at
the scale $Q = M_{\textrm{SUSY}} =
\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$
are the running $\overline{\textrm{DR}}$ stop masses evaluated at $Q =
M_{\textrm{SUSY}}$.
